\ripple: Differentiable Waveforms for Gravitational Wave Data Analysis

### 2022-12-17 src/tex/ms.tex
added scripts for figures

\newcommand{\jax}{\texttt{JAX}\xspace}
\newcommand{\ripple}{\texttt{ripple}\xspace}
\newcommand{\lalsuite}{\texttt{lalsuite}\xspace}
\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}

@@

\newcommand{\JHU}{\affiliation{William H. Miller III Department of Physics and Astronomy, Johns Hopkins University, Baltimore, Maryland 21218, USA}} 
\newcommand{\OKC}{\affiliation{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}} 
\newcommand{\NORDITA}{\affiliation{Nordic Institute for Theoretical Physics (NORDITA), 106 91 Stockholm, Sweden}}
\newcommand{\UdeM}{\affiliation{Département de Physique, Université de Montréal, 1375 Avenue Thérèse-Lavoie-Roux, Montréal, QC H2V 0B3, Canada}} 
\newcommand{\Mila}{\affiliation{Mila -- Quebec AI Institute, 6666 St-Urbain, \#200, Montreal, QC, H2S 3H1}} 
\newcommand{\CGP}{\affiliation{Center for Gravitational Physics, University of Texas at Austin, Austin, TX 78712, USA}}
\newcommand{\CCA}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}

% Author list
\author{Adam Coogan} \UdeM \Mila
\author{Thomas D.~P.~Edwards} \JHU \OKC \NORDITA
\author{Daniel Foreman-Mackey} \CCA
\author{Maximiliano Isi} \CCA
\author{Kelvin Lam} \CCA
\author{Kaze W.~K.~Wong} \CCA
\author{Aaron Zimmerman} \CGP

% Abstract with filler text
\begin{abstract}
    Since their discovery in 2015, gravitaitonal waves (GWs) have developed into a unique observational probe used throughout fundamental and astro-phyics.
    As more detections continue, data analysis taks are becoming increasingly computationally demanding.
    Here we propose the use of automatic differentiation through the programming framework \jax for accelerating a variety of these analsyis tasks.
    Firstly, we demonstrate that highly accurate waveforms (IMRPhenomD) can be written in \jax and demonstrate that the evaluation speed of the waveform (and its derivative) is not substantially slower than the \lalsuite implenetation in \CC.
    We then focus on three applications where speed and differentiable waveforms are essential.
    Firstly, we demonstrate that stochastic gradient descent can be used to optimize the $\sim 200$ coefficients that make up the waveform. 
    In particular, we demonstrate that the typical \textit{match} with numerical relativiely waveforms can be improved by more than 30\% without any additional overhead.
    Secondly, we show that Fisher forecasting calculations can be sped up by $\sim 100\times$ with no loss in accuracy.
    This increased speed makes population forecasting substantially simpler, as discussed extensively in \citep{Iacovelli:2022bbs, Iacovelli:2022mbg}.
    Finally, we show that gradient based samplers like Hamiltonian Monte Carlo can improve acceptance rates of XXX. 
    Since differentiable waveforms appear to have substatial advantages for a variety of tasks throughtout GW science (with little downside), we propose that waveform developers use \jax to build new waveforms moving forward.
    Our waveform code, called \ripple, can be found at XXX, and will continue to be updated with new waveforms as they are formulated.

\end{abstract}

@@

The discovery of gravitational waves (GWs) from inspiraling and mergering compact objects (COs) has revolutionized our understanding of both fundamental physics and astronomy.
Although the data volumes from GW detectors are relatively small, analyzing the data is a computationally demanding task.
In addition, this computational cost will substatially increase when next generation detectors come online.
The complexity begins before data gathering where one is required to generate banks of template waveforms which will be used for a matched-filter search~\citep{Owen:1998dk, Owen:1995tm}. 
Once potential candidates are found, parameter estimation (PE) is performed to extract the detailed source properties of each event.
In the most minimial example, this requires one to perform an MCMC on an 11 dimensional parameter space for each event.
Beyond this simple scenario, more complex waveform models with many additional parameters may be used to test for deviations from General Relativity.
Finally, using the results of the PE, population synthesis models are used to constrain the progenitors systems from which the black holes we see merging today began their journey~\citep{Wong:2022flg}.
All of these tasks require significant compute.
In this paper, we will argue that differentiable waveforms (and more generally differentiable pipelines) can play a significant role in alleaviating this computational demand.

Derivatives are ubiqitously useful throughout data analysis tasks.
For instance, during PE, derivative information can be used to guide the sampling towards regions with higher likelihood values (e.g. in Hamiltonian Monte Carlo~\citep{2017arXiv170102434B} or Gradient Decsent).
Gradients are particularly valuable for high dimensional spaces.
Unfortunately, in the field of GW data analysis, analytic derivatives of the necessary quantities (such as the likelihood) have historically been difficult to obtain.
Numerical derivatives also suffer from accuracy issues stemming from rounding or truncation errors.
However, recent progress in automatic differentiation (AD) has shown promise in allowing general, fast derivative calcuations for gravitational waveforms~\citep{Coogan:2022qxs}.

Automatic differentiaton is a family of methods used to compute machine precision derivatives with little computational ovearhead. 
AD's recent ascendance is primarily driven by its use in machine learning, particularly for derivative computations of neural networks which use gradient descent during training.
The core idea of AD is that each any mathematical function can be broken down into a small set of basic operations, each with a known differentiation rule.\footnote{
    Of course, non-differentiable functions exist and care must be taken when treating these special cases.
    }
The full derivative can then be contructed using the chain rule.
There are now a variety of AD implementations, most notably in deep learning frameworks such as \texttt{pytorch}~\citep{pytorch} and \texttt{tensorflow}~\citep{tensorflow2015-whitepaper}.
More general frameworks exist in \texttt{julia}~\citep{zygote, forwarddiff}, although \texttt{julia}'s limited support in GW physics precludes its general use.
Here we make use of \jax~\citep{jax2018github} due to its easy integration with \texttt{python} libraries, ability to just-in-time compile, and ability to utilize a variety of computational resources (i.e. CPUs and GPUs) without changes to the code.
\te{Do we need a longer intro to AD here?}

@@

In this paper we argue that differentiable waveforms will be a vital component for the future of GW data analysis.
In addition, we present \ripple, a small GW \texttt{python} package which can serve as a respository for differentiable implementations of some of the main waveforms currently used in LIGO and Virgo analyses. 
The remainder of this paper is structured as follows. 
In Sec.~\ref{sec:waveforms} we discuss the differentiable waveforms implemented in \ripple and perform some benchmarks to demonstrate the speed and accruacy of the derivative calculations. 
In Sec.~\ref{sec:applications} we discuss three distinct applications using differentiable waveforms. 
Firstly, we illustrate how the fit coefficients that form part of the IMRPhenom waveform models could be improved by high dimensional fitting, enabled by a differentiable waveform. 
Secondly, we implement differnetiable detector response functions and show that the speed of Fisher matrix calculations can be substantially improved using AD.
Finally, we run an illustrative injection example using Hamiltonian Monte Carlo to demonstrate that the acceptance rate of derivative samplers is substantially larger than traditional MCMCs.
The associated code can be found at \ripple~\citep{ripple}.

@@

Since the match is a measure of the difference between two waveforms, we can use it to demonstrate that the implementation of IMRPhenomD in \ripple accurately matches the \lalsuite implementation.
This is shown in Fig.~\ref{fig:match}, where we have calculated the $\mathrm{m}(h_{\theta_1}|h^{*}_{\theta_1})$ across the entire parameter space.\footnote{
    Specifically, we use 2000 points in the ranges $m_{1,2} = (1,50)\,M_{\odot}$ and $\chi_{1,2} = (-1,1)$. In addition, we evaulated the waveforms on a frequency grid from $32\,$Hz to $1024\,$Hz with frequency spacing $\Delta f = 0.0125 \,$Hz.
}
Here, $h_{\theta_1}$ corresponds to the \ripple waveform implementation and $h^{*}_{\theta_1}$ is the \lalsuite implementation. 
From Fig.~\ref{fig:match}, it is clear that the \ripple waveform matches with the \lalsuite waveform to numerical precision across the majority of the parameter space.
In particular, the blue points indicate parameter points where the $1-\mathrm{m}(h_{\theta_1}|h^{*}_{\theta_1})=0$ and would therefore not appear in the log scale colorbar.
\begin{figure}[t]
    \script{random_matches.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/random_matches.pdf}
        \caption{
            This plot should show a match colour bar.
        }
        \label{fig:match}
    \end{centering}
\end{figure}

@@

For maximum utility, a waveform needs to be fast to evaluate.
Fortunately, the Phenom IMR waveforms are constructed from simple closed form expressions which are therefore computationally efficient.
Since the \lalsuite implementation of IMRPhenomD is written in \texttt{C}, the evaluation of the waveform itself is faster than \ripple.
Benchmarking on a MacBook Pro with an M1 Max Apple Silicon processor, we find that a single waveform evaluation takes $\sim 2-5\,\mathrm{ms}$ for \lalsuite (interfaced with python) and $\sim 5-8\,\mathrm{ms}$ for \ripple.\footnote{
    For this benchmark we used $16\,\mathrm{Hz}$ and $1024\,\mathrm{Hz}$ for the lowest and highest frequencies respectively. 
    In addition, we used a frequency spacing of $0.0125\,\mathrm{Hz}$. 
    We performed this benchmark by evaluating the waveform $5000$ times and taking the average evaluation time.
}
Although slower on a CPU benchmark, a key advantage of using \jax is its native ability to run on a GPU.
Waveform calls can therefore be evaluated in parallel, massively improving performance.
Performing the same benchmark as above on a google colab notebook with the freely provided GPU, we find that on average waveform evaluations take $\sim 0.00005\mathrm{ms}$, nearly five orders of magnitude faster.
Generalizing \lalsuite waveforms to run on a GPU would be a significant undertaking.

@@

% \te{Kelvin should add stuff here}

@@

Every waveform approximant has some fitting coefficients that are fixed during
the process of calibrating the ansatz to NR waveforms. In the case of
IMRPhenomD, there are 209 fitting coefficients used to capture the separate
behavior of the amplitude and phase as a function of the mass ratio and spins.
The accuracy of fitting coefficients determines how well the approximant
reproduces the NR waveform. The inaccuracy in obtaining the fitting coefficients
means misrepresentation of the NR waveform, meaning the use of approximant will
introduce some systematic error in downstream data analysis task. For example, sufficiently large systematic errors 
would mean cause the recovered source parameters to be biased in the case of PE.

% Motivation
Previously in the construction of IMRPhenomD \citep{Khan:2015jqa}, waveform
coefficients are fitted in segments with amplitude and phase separated.
Individual segments are then connected together such that they are
differentiable. Using this method, the generated waveforms are expected to be
erroneous, as the connections would affect the originally fitted segments. The
coefficients of the current IMRPhenomD implementation are often hand tuned, and
are tuned in sub-sets of parameters instead of all together. This means the
tuning process ignore the correlation between different sub-sets of parameters,
so the best-fitted solution obtained by fitting sub-sets of parameters may not
be the global optimum compared to fitting all coefficients at once. Instead, by
jointly fitting all coefficients, the inaccuracy in fitting individual and
connecting different segments can be accounted during the optimization.
Therefore, there is potential accuracy one can gain by fitting all coefficients
at once.

\kw{Kelvin Fill}
Since we can differentiate through the entire waveform model against the fitting
parameters, one can use gradient descent to find the optimal fitting parameters
all at once instead of fitting them in subsets.

Here, we define the mismatch as
\begin{equation}
	\mathcal{M}=1-\mathrm{m}(h_{\mathrm{IMR}}|h_{\mathrm{NR}}), 
\end{equation} 
In our optimization procedure, we took NR waveforms stated in \citep{Khan:2015jqa} as training waveforms to fit waveform coefficients. 
By defining the mean of individual mismatches as the loss function for optimization, we can calibrate the set of coefficients using gradient descent.

\kw{Describe figure losscompare}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{../static/amplitude_loss.pdf}
\caption{Relative error of amplitudes of IMRPhenomD waveforms. Blue: Error
    between original Phenom waveform and NR waveform; Orange: Error between
    optimized waveform and NR waveform. \te{Can you add the data and the plotting script to the repo so that we can generate the plots using the same matplotlib settings?}}
\label{fig:loss_compare}
\end{figure}

\kw{Describe figure logdiff}
As demonstration, Fig. \ref{fig:loss_compare} shows the relative error of original and optimized waveform at different frequencies. 
In general, the error of optimized waveform is lower than that of the original waveform. 
The error in the early inspiral region is substantially decreased by half and other regions also show good improvement in accuracy.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{../static/MM_logdiff_q148_LIGO.pdf}
\caption{Distribution of the log difference of mismatches between optimized
    and original model. The mean shows a 36\% decrease in mismatch. }
\label{fig:MM_logdiff}
\end{figure}

Instead of examining individual waveforms, we investigate the general behavior of mismatches for a set of testing waveforms. 
In Fig. \ref{fig:logdiff_q148}, the distribution of log difference of mismatches are shown. 
The distribution mean lies towards the negative side, which indicates a improvement in the model. 
Despite there are testing waveforms that performed worse after optimization, it is verified that the ansatz of IMRPhenomD does not fit well with most of these waveforms, hence cannot be improved by optimization.  
Therefore, the new set of waveform coefficients can produce better waveforms for GW analyses.

\kw{Describe implication}
% Implication

This fine-tuning step is rather simple and general. Given a different
differentiable waveform model, one can apply the same procedure to optimize the
coefficients in the new model, i.e. applying gradient descent to optimize all
the fitting coefficients all together. This provides a more flexible control
during the calibration process, and potentially better final result.
With further modifications to the optimization process, the set of waveform
coefficients can be retuned to give almost an order-of-magnitude improvement.
\kw{Where is the proof for this?}

@@

\begin{figure}[t!]
    \script{sky_localization.py}
    \centering
    \includegraphics[width=\linewidth]{figures/sky_localization.pdf}
    \caption{
        This plot should show a sky localization error bar.
    }
    \label{fig:sky_localization}
\end{figure}

@@

\begin{table}[t]
    % \begin{center}
    \centering
    \begin{tabular}{c|c}
    \hline\hline
    $m_1, m_2$ & $\boldsymbol{U}[1.0, 50]\, \mathrm{M}_\odot$ \\ \hline
    $\chi_1, \chi_2$ & [-0.99, 0.99] \\ \hline
    $D$ & [500, 1000]\, Mpc \\ \hline
    $t_c$ & 0.0 \\ \hline
    $\phi_c$ & 0.0 \\ \hline
    Inclincation Angle, $\iota$ & $\boldsymbol{U}[0, 1]$ \\ \hline
    Polarization Angle, $\psi$ & $\boldsymbol{U}[0, 1]$ \\ \hline
    Right Ascension, $\delta$ & $\boldsymbol{U}[0, 1]$ \\ \hline
    Declination, $\alpha$ & $\boldsymbol{U}[0, 1]$ \\ 
    \hline\hline
    \end{tabular}
    % \end{center}
    \caption{Priors for the 11 dimensional parameter space used for the Fisher forecasting population analysis in Sec.~\ref{subsec:fisher}. 
    $\boldsymbol{U}$ indicates a uniform distribution between the two variables in the brackets.}
    \label{tab:priors}
\end{table}

@@

@@

\section{Acknowledgments}
This work was supported by collaborative visits funded by the Cosmology and Astroparticle Student and Postdoc Exchange Network (CASPEN). 
T.E. is supported by the Horizon Postdoctoral Fellowship.

>>>>

\newcommand{\jax}{\texttt{JAX}\xspace}
\newcommand{\ripple}{\texttt{ripple}\xspace}
\newcommand{\lalsuite}{\texttt{lalsuite}\xspace}

@@

\newcommand{\OKC}{\affiliation{The Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, SE-106 91 Stockholm, Sweden}} 
\newcommand{\NORDITA}{\affiliation{Nordic Institute for Theoretical Physics (NORDITA), 106 91 Stockholm, Sweden}}
\newcommand{\UdeM}{\affiliation{Département de Physique, Université de Montréal, 1375 Avenue Thérèse-Lavoie-Roux, Montréal, QC H2V 0B3, Canada}} 
\newcommand{\Mila}{\affiliation{Mila -- Quebec AI Institute, 6666 St-Urbain, \#200, Montreal, QC, H2S 3H1}} 
\newcommand{\CGP}{\affiliation{Center for Gravitational Physics, University of Texas at Austin, Austin, TX 78712, USA}}
\newcommand{\CCA}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
\newcommand{\CUHK}{\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, N.T., Hong Kong}}

% Author list
\author{Adam Coogan} \UdeM \Mila
\author{Thomas D.~P.~Edwards} \OKC \NORDITA
\author{Daniel Foreman-Mackey} \CCA
\author{Maximiliano Isi} \CCA
\author{Kelvin K.~H.~Lam} \CCA \CUHK
\author{Kaze W.~K.~Wong} \CCA
\author{Aaron Zimmerman} \CGP

% Abstract with filler text
\begin{abstract}
    Here we will discuss our implementation of differentiable waveforms and demonstrate their benefits for a variety of data analysis tasks.
\end{abstract}

@@

The discovery of gravitational waves (GWs) from inspiraling and mergering compact objects (COs) has revolutionized our understanding of both fundamental physics and astronomy.
Although the data volumes from GW detectors are relatively small, analyzing the data is a computationally demanding task.
In addition, this computational cost will substatially increase when next generation detectors come online.
The complexity begins before data gathering where one is required to generate banks of template waveforms which will be used for a matched-filter search~\citep{Owen:1998dk, Owen:1995tm}. 
Once potential candidates are found, parameter estimation (PE) is performed to extract the detailed source properties of each event.
In the most minimial example, this requires one to perform MCMC over a 15 dimensional parameter space for each event.
Beyond this simple scenario, more complex waveform models with many additional parameters may be used to test for deviations from General Relativity.
Finally, using the results of the PE, population synthesis models are used to constrain the progenitors systems from which the black holes we see merging today began their journey~\citep{Wong:2022flg}.
All of these tasks require significant compute.
In this paper, we will argue that differentiable waveforms (and more generally differentiable pipelines) can play a significant role in alleaviating this computational demand.

Derivatives are ubiqitously useful throughout data analysis tasks.
For instance, during PE, derivative information can be used to guide the sampling towards regions with higher likelihood values (e.g. in Hamiltonian Monte Carlo~\citep{2017arXiv170102434B} or Gradient Decsent).
This use of gradients is particularly useful for high dimensional spaces.
Unfortunately, in the field of GW data analysis, analytic derivatives of the necessary quantities (such as the likelihood) have historically been difficult to obtain.
Numerical derivatives also suffer from accuracy issues stemming from rounding or truncation errors.
However, recent progress in automatic differentiation (AD) has shown promise in allowing general, fast derivative calcuations for gravitational waveforms~\citep{Coogan:2022qxs}.

Automatic differentiaton is a family of methods used to compute machine precision derivatives with little computational ovearhead. 
AD's recent ascendance is primarily driven by its use in machine learning, particularly for derivative computations of neural networks which use gradient descent during training.
The core idea of AD is that each any mathematical function can be broken down into a small set of basic operations, each with a known differentiation rule.\footnote{
    Of course, non-differentiable functions exist and care must be taken when treating these special cases.
    }
The full derivative can then be contructed using the chain rule.
There are now a variety of AD implementations, most notably in deep learning frameworks such as \texttt{pytorch}~\citep{pytorch} and \texttt{tensorflow}~\citep{tensorflow2015-whitepaper}.
More general frameworks exist in \texttt{julia}~\citep{zygote, forwarddiff}, although \texttt{julia}'s limited use in GW physics precludes its general use.
Here we make use of \jax~\citep{jax2018github} due to its easy integration with \texttt{python} libraries, ability to just-in-time compile, and ability to utilize a variety of computational resources (i.e. CPUs and GPUs) without changes to the code.
\te{Do we need a longer intro to AD here?}

@@

In this paper we argue that differentiable waveforms will be a vital component for the future of GW data analysis.
In addition, we present \ripple, a small GW \texttt{python} package containing differentiable implementations of some of the main waveforms currently used in LIGO and Virgo analyses. 
The remainder of this paper is structured as follows. 
In Sec.~\ref{sec:waveforms} we discuss the differentiable waveforms implemented in \ripple and perform some benchmarks to demonstrate the speed and accruacy of the derivative calculations. 
In Sec.~\ref{sec:applications} we discuss four distinct applications using differentiable waveforms. 
First, a simple demonstration of accelerated effectualness calculations using gradient descent.
Second, we show that Fisher analyses are quickly enabled using AD.
Third, we implement differnetiable detector response functions and show an illutrative example using Hamiltonian Monte Carlo.
Finally, we illustrate how the fit coefficients that form part of the IMRPhenom waveform models could be improved by high dimensional fitting, enabled by a differentiable waveform. 
The associated code can be found at \ripple~\citep{ripple}.

@@

Since the match is a measure of the difference between two waveforms, we can use it to demonstrate that the implementation of IMRPhenomD in \ripple accurately matches the \lalsuite implementation.
This is shown in Fig.~\ref{fig:match}, where we have calculated the $\mathrm{m}(h_{\theta_1}|h^{*}_{\theta_1})$ across the entire parameter space.\footnote{
    Specifically, we use 2000 points in the ranges $m_{1,2} = (1,50)\,M_{\odot}$ and $\chi_{1,2} = (-1,1)$. In addition, we evaulated the waveforms on a frequency grid from $32\,$Hz to $1024\,$Hz with frequency spacing $\Delta f = 0.0125 \,$Hz.
}
Here, $h_{\theta_1}$ corresponds to the \ripple waveform implementation and $h^{*}_{\theta_1}$ is the \lalsuite implementation. 
From Fig.~\ref{fig:match}, it is clear that the \ripple waveform matches with the \lalsuite waveform to numerical precision across the majority of the parameter space.
In particular, the blue points indicate parameter points where the $1-\mathrm{m}(h_{\theta_1}|h^{*}_{\theta_1})=0$ and would therefore not appear in the log scale colorbar.
\begin{figure}[ht!]
    \script{random_matches.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/random_matches.pdf}
        \caption{
            This plot should show a match colour bar.
        }
        \label{fig:match}
    \end{centering}
\end{figure}

@@

For maximum utility, a waveform needs to be fast to evaluate.
Fortunately, the Phenom IMR waveforms are constructed from simple closed form expressions which are therefore computationally efficient.
Since the \lalsuite implementation of IMRPhenomD is written in \texttt{C}, the evaluation of the waveform itself is faster than \ripple.
Benchmarking on a MacBook Pro with an M1 Max Apple Silicon processor, we find that a single waveform evaluation takes $\sim 4\,\mathrm{ms}$ for \lalsuite (interfaced with python) and $\sim 10\,\mathrm{s}$ for \ripple.\footnote{
    For this benchmark we used $16\,\mathrm{Hz}$ and $1024\,\mathrm{Hz}$ for the lowest and highest frequencies respectively. 
    In addition, we used a frequency spacing of $0.0125\,\mathrm{Hz}$. 
    We performed this benchmark by evaluating the waveform $5000$ times and taking the average evaluation time.
}
Although slower on a CPU benchmark, a key advantage of using \jax is its native ability to run on a GPU.
Waveform calls can therefore be evaluated in parallel, massively improving performance. \te{I couldn't get these to be similar on google colab. @Kaze any ideas how to make this fast?}
Generalizing \lalsuite waveforms to run on a GPU would be a significant undertaking.

@@

@@

Every waveform approximants have some fitting coefficients that are fixed during
the process of calibrating the ansatz to NR waveforms. In the case of
IMRPhenomD, there are 209 fitting coefficients used to capture the separate
behavior of the amplitude and phase as a function of the mass ratio and spins.
The accuracy of fitting coefficients determines how well the approximant
reproduces the NR waveform. The inaccuracy in obtaining the fitting coefficients
means misrepresentation of the NR waveform, meaning the use of approximant will
introduce some systematic error in downstream data analysis task. For example,
this would mean the recovered source parameters will be biased in the case of PE.

Previously in the construction of IMRPhenomD \citep{Khan:2015jqa}, waveform
coefficients are fitted in segments with amplitude and phase separated.
Individual segments are then connected together such that they are
differentiable. Using this method, the generated waveforms are expected to be
erroneous, as the connections would affect the originally fitted segments. The
coefficients of the current IMRPhenomD implementation are often hand tuned, and
are tuned in sub-sets of parameters instead of all together. This means the
tuning process ignore the correlation between different sub-sets of parameters,
so the best-fitted solution obtained by fitting sub-sets of parameters may not
be the global optimum compared to fitting all coefficients at once. Instead, by
jointly fitting all coefficients, the inaccuracy in fitting individual and
connecting different segments can be accounted during the optimization.
Therefore, there is potential accuracy one can gain by fitting all coefficients
at once.

Since we can differentiate through the entire waveform model against the fitting
parameters, one can use gradient descent to find the optimal fitting parameters
all at once instead of fitting them in subsets.

Here, we define the mismatch and loss function as
\begin{equation}
	\mathcal{M}&=1-\mathrm{m}(h_{\mathrm{IMR}}|h_{\mathrm{NR}}).
\end{equation}
In the optimization procedure, we took NR waveforms $h_{\mathrm{NR}}$ stated in \citep{Khan:2015jqa} as training waveforms to fit waveform coefficients.  
Then, combining individual mismatches, a loss function is constructed as follows:
\begin{equation}
	\mathcal{L}&=\frac{1}{N}\sum_{i=1}^{N}\mathcal{M}_i, 
\end{equation}
where $\mathcal{M}_i$ is the mismatch of individual training waveforms and $N$ is the total number of training waveforms used in optimization. 
Then, one can proceed to perform optimization on waveform coefficients.

\begin{figure}[t]
	\script{loss_compare.py}
	\begin{centering}
		\includegraphics[width=\linewidth]{figures/loss_compare.pdf}
		\caption{
			Relative error of IMRPhenomD waveform amplitudes. Blue: Error
			between original Phenom waveform and NR waveform; Orange: Error between
			optimized waveform and NR waveform.
		}
		\label{fig:loss_compare}
	\end{centering}
\end{figure}

As demonstration, Fig. \ref{fig:loss_compare} shows the relative error of original and optimized waveform at different frequencies. 
In general, the error of optimized waveform is lower than that of the original waveform. 
The error in the early inspiral region is decreased by half while other regions also show good improvement in accuracy.

\begin{figure}[t]
	\script{mismatch_hist.py}
	\begin{centering}
		\includegraphics[width=\linewidth]{figures/mismatch_hist.pdf}
		\caption{
			Distributions of log mismatches of training waveforms. Blue: Original waveforms; Orange: Optimized waveforms.
		}
		\label{fig:mismatch_hist}
	\end{centering}
\end{figure}

Instead of examining individual waveforms, we investigate the general behavior of mismatches for a set of testing waveforms. 
In Fig. \ref{fig:mismatch_hist}, the distribution of log mismatches are shown. 
The entire optimized waveform distribution is slightly shifted to the negative side. 
Waveforms originally with moderate mismatch moved down to the left end of the distribution. 
Both indicate that the optimization procedure improved the model. 
Therefore, the new set of waveform coefficients can produce better waveforms for GW analyses.

This fine-tuning step is rather simple and general. Given a different
differentiable waveform model, one can apply the same procedure to optimize the
coefficients in a new model, i.e. applying gradient descent to optimize all
the fitting coefficients all together. This provides a more flexible control
during the calibration process, and potentially better final result.
With further modifications to the optimization process, the set of waveform
coefficients can be retuned to give almost $50\%$ improvement in mismatch for well-behaved waveforms.
%\kw{Where is the proof for this?}

@@

@@

\begin{table}[t!]
    % \begin{center}
    \centering
    \begin{tabular}{c|c}
    \hline\hline
    $m_1, m_2$ & $\boldsymbol{U}[1.0, 50]\, \mathrm{M}_\odot$ \\ \hline
    $\chi_1, \chi_2$ & [-0.99, 0.99] \\ \hline
    $D$ & [500, 1000]\, Mpc \\ \hline
    $t_c$ & 0.0 \\ \hline
    $\phi_c$ & 0.0 \\ \hline
    Inclincation Angle, $\iota$ & $\boldsymbol{U}[0, 1]$ \\ \hline
    Polarization Angle, $\psi$ & $\boldsymbol{U}[0, 1]$ \\ \hline
    Right Ascension, $\delta$ & $\boldsymbol{U}[0, 1]$ \\ \hline
    Declination, $\alpha$ & $\boldsymbol{U}[0, 1]$ \\ 
    \hline\hline
    \end{tabular}
    % \end{center}
    \caption{Priors for the 11 dimensional parameter space used for the Fisher forecasting population analysis in Sec.~\ref{subsec:fisher}. 
    $\boldsymbol{U}$ indicates a uniform distribution between the two variables in the brackets.}
    \label{tab:priors}
\end{table}

@@

\begin{figure}[ht!]
    \script{sky_localization.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/sky_localization.pdf}
        \caption{
            This plot should show a sky localization error bar.
        }
        \label{fig:sky_localization}
    \end{centering}
\end{figure}

@@

\section{Acknowledgments}
This work was supported by collaborative visits funded by the Cosmology and Astroparticle Student and Postdoc Exchange Network (CASPEN).
